Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-25 paper.qmd
Updated Five Gau UMAP projections

## First projection
projection <- cbind(
    c(0.36638,0.31467,0.63030,0.24465),
    c(0.14672,-0.73585,0.33246,-0.12977))

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.45, -0.3)) +
  ylim(c(-0.05, 0.2)) +
  interior_annotation("c", 
                      position = c(0.08, 0.9),
                      cex = 3)

## Second projection
projection <- cbind(
    c(-0.70767,-0.09092,0.15109,-0.39811),
    c(-0.01312,-0.61805,0.44418,0.33305))

@@

five_gau_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.13, 0.13)) +
  ylim(c(-0.13, 0.13)) +
  interior_annotation("d", 
                      position = c(0.08, 0.9),
                      cex = 3)

@@

```{r}
#| label: fig-five-gau-projs
#| fig-height: 12
#| fig-width: 8
#| fig-pos: H
#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"

>>>>

## First projection
projection <- cbind(
  c(0.02743,0.21427,0.62893,-0.49816),
  c(-0.70709,0.42156,-0.02593,0.10965))

@@

# axes_obj <- gen_axes(
#   data = scaled_gau_data, 
#   proj_data = projected_df, 
#   limits = 1)
# 
# axes <- axes_obj$axes
# circle <- axes_obj$circle

limits <- 1
projected_df_n <- projected_df |> select(-ID)

rng <- range(projected_df_n)
projected_df_n <- projected_df_n/max(abs(rng))
colnames(projected_df_n) <- c("P1", "P2")
projected_df_n <- data.frame(projected_df_n)
obs_labels <- as.character(1:nrow(scaled_gau_data))

axis_scale <- limits/6
# axis_pos <- -2/3 * limits
axis_pos <- -1

axis_labels <- colnames(scaled_gau_data)
rownames(axes) <- axis_labels

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5,
    alpha = 0.2) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-1.25, 0.1)) +
  ylim(c(-1.25, -0.5)) +
  interior_annotation("c", 
                      position = c(0.08, 0.9),
                      cex = 3)

## Second projection
projection <- cbind(
  c(0.36030,0.00630,0.66545,-0.34307),
  c(-0.01861,0.14409,-0.36798,-0.73066))

@@

# axes_obj <- gen_axes(
#   data = scaled_gau_data, 
#   proj_data = projected_df, 
#   limits = 1)
# 
# axes <- axes_obj$axes
# circle <- axes_obj$circle

limits <- 0.3
projected_df_n <- projected_df |> select(-ID)

rng <- range(projected_df_n)
projected_df_n <- projected_df_n/max(abs(rng))
colnames(projected_df_n) <- c("P1", "P2")
projected_df_n <- data.frame(projected_df_n)
obs_labels <- as.character(1:nrow(scaled_gau_data))

axis_scale <- limits/6
# axis_pos <- -2/3 * limits
axis_pos <- 0.08

axis_labels <- colnames(scaled_gau_data)
rownames(axes) <- axis_labels

five_gau_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#33a02c") +
  geom_point(
    size = 0.5, 
    alpha = 0.3) +
  scale_color_manual(values = c("#000000", "#33a02c")) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(0, 0.4)) +
  ylim(c(0, 0.3)) +
  interior_annotation("d", 
                      position = c(0.08, 0.9),
                      cex = 3)

@@

```{r}
#| label: fig-five-gau-projs
#| fig-height: 15
#| fig-width: 10
#| fig-pos: H
#| fig-cap: "(a) tSNE, (b) UMAP, and (c) PaCMAP"
